% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmtoplas]{acmsmall} % Aptara syntax
\usepackage{etex}				% extend latex, especially the number of registers

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} % ou \usepackage[OT1]{fontenc}
\usepackage[frenchb]{babel} % optionnel

\usepackage[dvipsnames]{xcolor}
\usepackage{pgfplots}			% plotting graphs
\pgfplotsset{compat=1.12}
\usetikzlibrary{patterns}

\usepackage{bussproofs}

\graphicspath{{./images/}}
\usepackage[caption=false,font=footnotesize]{subfig}

\usepackage{multirow}
\usepackage{multicol}
\usepackage{array}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{listings}
\lstdefinestyle{full}{
	language=C,
	numbers=left,
	captionpos=b,
	breaklines=true,
	frame=single,
	basicstyle=\footnotesize,
	tabsize=2,
	escapeinside={(*}{*)},
	moredelim=**[is][\color{Red}]{@}{@},
	moredelim=**[is][\color{blue}]{@@}{@@},
	moredelim=**[is][\color{Green}]{@@@}{@@@},
	morekeywords={par, weak, abort, when, input, output, shared, pause, combine, all, new, mod, with, status, preempt, copy, nop
		  life, mission, noncritical, Hz,
		  EOT, PAR,
		  ||, signal, module, emit, present, then, end, immediate}
}

\lstdefinestyle{fulltight}{
	language=C,
	numbers=left,
	captionpos=b,
	breaklines=true,
	frame=single,
	basicstyle=\scriptsize\sffamily,
	tabsize=2,
	escapeinside={(*}{*)},
	columns=flexible,
	xleftmargin=20pt,
	moredelim=**[is][\color{Red}]{@@}{@@},
	morekeywords={par, weak, abort, when, input, output, shared, pause, combine, all, new, mod, with, status, preempt, copy, nop
		  life, mission, noncritical, Hz,
		  EOT, PAR,
		  ||, signal, module, emit, present, then, end, immediate}
}

\lstdefinestyle{snippet}{
	language=C,
	basicstyle=\small,
	frame=none,
	captionpos=b,
	numbers=none,
	xleftmargin=16pt,
	breaklines=true,
	tabsize=2,
	escapeinside={(*}{*)},
	morekeywords={par, weak, abort, when, input, output, shared, pause, combine, all, new, mod, with, 
		  life, mission, noncritical, Hz,
		  EOT, PAR,
		  ||, signal, module, emit, present, then, end, immediate}
}

\newcommand{\StateP}{\ensuremath{S}}
\newcommand{\Abort}{\ensuremath{A}}
\newcommand{\abort}{\ensuremath{a}}
\newcommand{\Environment}{\ensuremath{E}}
\newcommand{\Input}{\ensuremath{I}}

\newcommand{\Id}{\ensuremath{\mathit{Id}}}
\newcommand{\id}{\ensuremath{\mathit{id}}}
\newcommand{\Global}{\ensuremath{\mathcal{G}}}
\newcommand{\Store}{\ensuremath{Store}}
\newcommand{\store}{\ensuremath{store}}
\newcommand{\Var}{\ensuremath{Var}}
\newcommand{\var}{\ensuremath{var}}
\newcommand{\Val}{\ensuremath{Val}}
\newcommand{\val}{\ensuremath{v}}
\newcommand{\Sts}{\ensuremath{Sts}}
\newcommand{\sts}{\ensuremath{sts}}
\newcommand{\expression}{\ensuremath{exp}}
\newcommand{\imm}{\ensuremath{imm}}

\newcommand{\thread}[1]{\ensuremath{\mathit{t}_{#1}}}
\newcommand{\Thread}{\ensuremath{T}}
\newcommand{\body}[1]{\ensuremath{\mathit{f}_{#1}}}

\DeclareMathOperator{\GetParent}{\textsc{GetParent}}
\DeclareMathOperator{\GetPolicy}{\textsc{GetPolicy}}
\DeclareMathOperator{\GetShared}{\textsc{GetShared}}
\DeclareMathOperator{\GetCombine}{\textsc{GetCombine}}
\DeclareMathOperator{\GetExp}{\textsc{GetExp}}

\DeclareMathOperator{\Eval}{\textsc{Eval}}
\DeclareMathOperator{\GetVal}{\textsc{GetVal}}
\DeclareMathOperator{\Copy}{\textsc{Copy}}
\DeclareMathOperator{\Combine}{\textsc{Combine}}

\newdef{hypothesis}[theorem]{Hypothesis}

\newcommand{\change}[1]{{\color{red} #1}}
%\newcommand{\change}[1]{#1}

\newcommand{\pretc}{\mbox{PRET-C}}
\newcommand{\synchronousc}{\mbox{Synchronous C}}

\hyphenation{ForeC}


% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled,linesnumbered]{algorithm2e}
\DontPrintSemicolon
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{0}
\acmNumber{0}
\acmArticle{00}
\acmYear{0000}
\acmMonth{0}

% Copyright
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{0000001.0000001}

%ISSN
\issn{1234-56789}

% Document starts
\begin{document}

% Page heads
\markboth{E. Yip et al.}{Synchronous Deterministic Parallel Programming for Multicores with ForeC}

% Title portion
\title{Synchronous Deterministic Parallel Programming for Multicores with ForeC}
\author{
	EUGENE YIP
	\affil{Department of ECE, The University of Auckland, New Zealand}
	PARTHA S. ROOP
	\affil{Department of ECE, The University of Auckland, New Zealand}
	ALAIN GIRAULT
	\affil{Inria, France. Universit{\'e} Grenoble Alpes, Lab. LIG, Grenoble, France. CNRS, Lab. LIG, F-38000 Grenoble, France}
	MORTEZA BIGLARI-ABHARI
	\affil{Department of ECE, The University of Auckland, New Zealand}
}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}
	Cyber-physical systems (CPSs) are embedded systems that are tightly 
	integrated with their physical environment. The correctness of a 
	CPS depends on the output of its computations and on the timeliness
	of completing the computations. The increasing use of high-performing and 
	low-power multi-core processors in embedded systems is pushing 
	embedded programmers to 
	be parallel programming experts. Parallel programming is challenging 
	because of the skills, experiences, and knowledge needed to avoid 
	common parallel programming traps and pitfalls.
	This paper proposes
	the ForeC language for the deterministic, parallel, and reactive 
	programming of embedded multi-cores. The synchronous semantics of ForeC 
	is designed to greatly simplify the understanding and debugging
	of parallel programs. ForeC allows programmers to express many 
	forms of parallel patterns while ensuring that ForeC programs 
	can be compiled efficiently for parallel execution and be 
	amenable to static timing analysis. ForeC's main innovation 
	is its shared variable semantics that provides 
	thread isolation and deterministic thread communication. All 
	ForeC programs are correct by construction and deadlock-free 
	because mutual exclusion constructs are not needed.
	Through benchmarking, we demonstrate that ForeC can achieve better
	parallel performance than Esterel, a widely used synchronous language
	for concurrent safety-critical systems, and OpenMP, a popular 
	desktop solution for parallel programming. We demonstrate
	that the worst-case execution time of ForeC programs can be estimated
	to a high degree of precision.
\end{abstract}


%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
	<ccs2012>
	<concept>
		<concept_id>10010147.10010169.10010175</concept_id>
		<concept_desc>Computing methodologies~Parallel programming languages</concept_desc>
		<concept_significance>500</concept_significance>
	</concept>
	<concept>
		<concept_id>10011007.10011006.10011039.10011311</concept_id>
		<concept_desc>Software and its engineering~Semantics</concept_desc>
		<concept_significance>500</concept_significance>
	</concept>
	<concept>
		<concept_id>10011007.10011006.10011041.10011047</concept_id>
		<concept_desc>Software and its engineering~Source code generation</concept_desc>
		<concept_significance>300</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Parallel programming languages}
\ccsdesc[500]{Software and its engineering~Semantics}
\ccsdesc[300]{Software and its engineering~Source code generation}

%
% End generated code
%

% We no longer use \terms command
%\terms{Design, Algorithms, Performance}

\keywords{programming language, semantics, parallelism, synchronous, determinism, reactive, multi-core, worst-case execution time, code generation}

\acmformat{Eugene Yip, Partha S. Roop, Alain Girault, and Morteza Biglari-Abhari, 2016. 
Synchronous Deterministic Parallel Programming for Multicores with ForeC.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
	This work was supported in part by the RIPPES INRIA International Lab. \newline
	E. Yip, (Current address) Software Technologies Research Group, University of Bamberg, 96045 Germany.
\end{bottomstuff}

\maketitle

%\section{Introduction}
\input{./sections/introduction}

%\section{Related Work}
\input{./sections/literature}

%\section{Architecture of the Multicore Processor}
\input{./sections/architecture}

%\section{The ForeC Language}
\input{./sections/forec}

%\section{The Formal Semantics of ForeC}
\input{./sections/forec_semantics}
\input{./sections/forec_illustration1}
\input{./sections/forec_illustration2}

%\section{Proofs}
\input{./sections/forec_proofs1}
%\input{./sections/forec_proofs2}

\input{./sections/forec_comparison}

%\section{Compilation}
\input{./sections/forec_compiling}

%\section{Benchmarking}
\input{./sections/forec_benchmarking1}

%\section{Conclusions and Future Work}
\input{./sections/conclusion}

% Appendix
\appendix
\include{./sections/appendix}

\appendixhead{YIP}

% Acknowledgments
\begin{acks}
	We wish to acknowledge Pascal Fradet for his discussions on
	an earlier version of the {ForeC} semantics, and
	Jean-Bernard Stefani for his discussions on
	program equivalence in the context of {ForeC}.
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{./references/GeneralHardwareSoftware,./references/Wireless,./references/PRET}
% Sample .bib file with references that match those in
% the 'Specifications Document (V1.5)' as well containing
% 'legacy' bibs and bibs with 'alternate codings'.
% Gerry Murray - March 2012

% History dates
\received{July 2016}{July 2016}{July 2016}

% Electronic Appendix
%\elecappendix

%\medskip

%\section{This is an example of Appendix section head}

%Channel-switching time is measured as the time length it takes for
%motes to successfully switch from one channel to another. This
%parameter impacts the maximum network throughput, because motes
%cannot receive or send any packet during this period of time, and it
%also affects the efficiency of toggle snooping in MMSN, where motes
%need to sense through channels rapidly.

\end{document}
% End of v2-acmsmall-sample.tex (March 2012) - Gerry Murray, ACM


